---
title: "Map Visualizations and Neighorhood Change"
author: Prof. Anthony Howell
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: no
    css: 'lab-instructions.css'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set( echo=TRUE, message=F, warning=F, eval=T, fig.width=10 )
```



# Part 1: How not to lie with choropleth maps

## Grab Census Data (review from Lab 2)


- Using ```tidycensus```, load in variables on median household income and median housing value from the ```acs5``` 2013-2017. 

```{r}
library(tidycensus)
library(tidyverse)
library(viridis)
library(plyr)
library(gtools)
```

```{r, echo=F}
census_api_key("8eab9b16f44cb26460ecbde164482194b7052772")
```

```{r, eval=F}
census_api_key("Your Key here")
```


```{r}

Var <- c("B19013_001","B25077_001")
## c(Median household Income, Median Housing Value)

CenDF <- get_acs(geography = "county",
                   variables = Var,
                   year = 2017,
                   survey = "acs5",
                   geometry = TRUE,
                   shift_geo = TRUE) 
```


## Manipulate Data

- Create new variable for the housing price to income ratio. 

```{r}

CenDF <-
  CenDF %>% 
    mutate(variable=case_when( 
      variable=="B19013_001" ~ "HHIncome",
      variable=="B25077_001" ~ "HouseValue")) %>%
    select(-moe) %>%  
    spread(variable, estimate) %>%  #Spread moves rows into columns
    mutate(HHInc_HousePrice_Ratio=round(HouseValue/HHIncome,2)) 
```




## Map Data

- Rely on ggplot and ```scale_fill_viridis``` to create natural breaks for color sequencing and to mapHouse-Price-to-Income Ratio

```{r, out.width = '50%', fig.align='center', fig.width=9, fig.height=9}
Plot1 <-  ggplot(CenDF) +
    geom_sf(aes(fill = HHInc_HousePrice_Ratio), color=NA) +
    coord_sf(datum=NA) +
    labs(title = "House-Price-to-Income Ratio \n R Built-in Color Cut",
         caption = "Source: ACS 5-year, 2013-2017",
         fill = "Price-Income Ratio") +
    scale_fill_viridis(direction=-1)
Plot1
```


## Color Scales in R

R has a variety of functions that allow the user to create color scales easily. These functions generally require you to specify a color on each end of the spectrum and they will interpolate the values between based upon how many levels you desire.

## Importance

We select colors to communicate information about our data. If we are using a continuous variable the most basic decision is whether we want to represent it as positive and negative deviations from the average (a divergent scale), or as a continuum of low to high values (a sequential scale). If we have categorical data, it is generally visualized through different colors representing each group (a qualitative scale).



## Color Schemes

```{r out.width = '50%', fig.align='center',fig.width=5,fig.height=5}
plot( 1:7, rep(1,7), ylim=c(-0.5,3.5), xlim=c(0,12), yaxt="n", xaxt="n", bty="n", xlab="", ylab=""  )

color.function <- colorRampPalette( c("gray80","darkred") )
col.ramp <- color.function( 7 ) # number of groups you desire
points( 1:7, rep(3,7), pch=15, cex=8, col=col.ramp )

color.function <- colorRampPalette( c("darkred","gray80","steelblue") )
col.ramp <- color.function( 7 ) # number of groups you desire
points( 1:7, rep(2,7), pch=15, cex=8, col=col.ramp )

color.function <- colorRampPalette( c("gray80","black") )
col.ramp <- color.function( 7 ) # number of groups you desire
points( 1:7, rep(1,7), pch=15, cex=8, col=col.ramp )

text( 8, 3, "Sequential", pos=4 )
text( 8, 2, "Divergent", pos=4 )
text( 8, 1, "Grayscale", pos=4 )
```


## Alternative Map: 2 Color Groups

- Transform ```HHInc_HousePrice_Ratio``` from continuous variable to factor variable with 2 levels

```{r, out.width = '50%', fig.align='center',fig.width=9,fig.height=9}
## Convert Integer into factor variable
CenDF$fill_factor <- quantcut(CenDF$HHInc_HousePrice_Ratio, q = seq(0, 1, by = 0.5))
CenDF$fill_factor = mapvalues(CenDF$fill_factor, from = levels(CenDF$fill_factor),
                        to = c("low","high"))


## Assign 2 colors
col.ramp <- viridis(n = 2) # number of groups you desire
```



```{r, out.width = '50%', fig.align='center',fig.width=9,fig.height=9}

Plot2<-ggplot(CenDF) +
  geom_sf(aes(fill = fill_factor), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "House-Price-to-Income Ratio",
       caption = "Source: ACS 5-year, 2013-2017",
       fill = "Price-Income Ratio") +
  scale_fill_manual("Price-Income Ratio",values =  col.ramp) 
Plot2
```


## Terciles - 3 Color Groups

- Transform ```HHInc_HousePrice_Ratio``` from continuous variable to factor variable with 3 levels

```{r, out.width = '50%', fig.align='center',fig.width=9,fig.height=9}
## Convert Integer into factor variable
CenDF$fill_factor <- quantcut(CenDF$HHInc_HousePrice_Ratio, q = seq(0, 1, by = .33))
CenDF$fill_factor = mapvalues(CenDF$fill_factor, from = levels(CenDF$fill_factor),
                        to = c("low","med", "high"))


## Assign 3 colors
col.ramp <- viridis(n = 3) # number of groups you desire
```



```{r, out.width = '50%', fig.align='center',fig.width=9,fig.height=9}

Plot3<-ggplot(CenDF) +
  geom_sf(aes(fill = fill_factor), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "House-Price-to-Income Ratio",
       caption = "Source: ACS 5-year, 2013-2017",
       fill = "Price-Income Ratio") +
  scale_fill_manual("Price-Income Ratio \n (Terciles)",values =  col.ramp) 
Plot3
```




## Quintiles - 5 Color Groups

- Transform ```HHInc_HousePrice_Ratio``` from continuous variable to factor variable with 5 levels

```{r}
## Convert Integer into factor variable
CenDF$fill_factor <- quantcut(CenDF$HHInc_HousePrice_Ratio, q = seq(0, 1, by = .2))
CenDF$fill_factor = mapvalues(CenDF$fill_factor, from = levels(CenDF$fill_factor),
                        to = c("1","2","3", "4","5"))


## Assign 5 colors
col.ramp <- viridis(n = 5) # number of groups you desire
```



```{r, out.width = '50%', fig.align='center',fig.width=9,fig.height=9}

Plot4<-ggplot(CenDF) +
  geom_sf(aes(fill = fill_factor), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "House-Price-to-Income Ratio",
       caption = "Source: ACS 5-year, 2013-2017",
       fill = "Price-Income Ratio") +
  scale_fill_manual("Price-Income Ratio  \n (Quintiles)",values =  col.ramp) 
Plot4
```


## Unequal Intervals

- Transform ```HHInc_HousePrice_Ratio``` from continuous variable to factor variable with 6 levels

```{r}
## Convert Integer into factor variable
CenDF$fill_factor <- quantcut(CenDF$HHInc_HousePrice_Ratio, q = c(0,.1,.25,.5,.75,.9,1))
#CenDF$fill_factor = mapvalues(CenDF$fill_factor, from = levels(CenDF$fill_factor),
 #                       to = c("< .1",".1-.25",".25-.5", ".5-.75",".75-.9","> .9"))


## Assign 6 colors
col.ramp <- viridis(n = 6) # number of groups you desire
```



```{r, out.width = '50%', fig.align='center',fig.width=9,fig.height=9}

Plot5<-ggplot(CenDF) +
  geom_sf(aes(fill = fill_factor), color=NA) +
  coord_sf(datum=NA) +
  labs(title = "House-Price-to-Income Ratio",
       caption = "Source: ACS 5-year, 2013-2017",
       fill = "Price-Income Ratio") +
  scale_fill_manual("Price-Income Ratio",values =  col.ramp) 
Plot5
```


## Compare Maps

```{r, out.width = '100%', fig.align='center',fig.width=9,fig.height=9}
library(gridExtra)
grid.arrange(Plot1,Plot3,Plot4,Plot5, nrow=2)
```





# Example of Skew Correction

```{r}
ggplot(CenDF) +
    geom_sf(aes(fill = HHInc_HousePrice_Ratio), color=NA) +
    coord_sf(datum=NA) +
    labs(title = "House-Price-to-Income Ratio",
         caption = "Source: ACS 5-year, 2013-2017",
         fill = "Price-Income Ratio") +
    scale_fill_viridis(direction=-1)
```


## Revising the Scale for Skewed Data

Our price ratio is skewed because of the long right tail: 

```{r}
hist( CenDF$HHInc_HousePrice_Ratio, 
      breaks=50, col="darkgray" )
```

As a result, most of our data will be crammed into the bottom half of our scale, making the colors a lot harder to interpret. 

We can consider ways to transform the scale before visualization. 

Logging values can pull-in outliers: 


```{r}
hist( log( CenDF$HHInc_HousePrice_Ratio + 1 ), 
      breaks=50, col="darkgray" )
```

We can also top-code scales to remove the long tail: 

```{r}
# top-coding 
CenDF$inc.home.ratio <- CenDF$HHInc_HousePrice_Ratio
CenDF$inc.home.ratio[ CenDF$inc.home.ratio > 5 ] <- 5

hist( CenDF$inc.home.ratio, 
      breaks=50, col="darkgray" )
```


Let's also select a divergent scale to better differentiate high and low values (a neutral color like gray represents average values). 

The **pals** package has some crisp color scales. 

```{r}
### devtools::install_github("kwstat/pals")   
library( pals )   
pal.map( ocean.balance, n=13, main="Example Divergent Colorscale")
```


Let's see how this changes our interpretation of the data: 


```{r}
## REGULAR DIVERGENT SCALE WHEN DATA IS SKEWED
ggplot(CenDF) +
    geom_sf(aes(fill = HHInc_HousePrice_Ratio), color=NA) +
    coord_sf( datum=NA ) +
    labs( title = "House-Price-to-Income Ratio",
          caption = "Source: ACS 5-year, 2013-2017",
          fill = "Price-Income Ratio" ) +
    scale_fill_gradientn( colours=ocean.balance(10), guide = "colourbar" )


## TOP-CODED SCALE - MAX IS RATIO OF 6

CenDF$inc.home.ratio <- CenDF$HHInc_HousePrice_Ratio
CenDF$inc.home.ratio[ CenDF$inc.home.ratio > 6 ] <- 6

ggplot(CenDF) +
    geom_sf(aes(fill = inc.home.ratio), color=NA) +
    coord_sf( datum=NA ) +
    labs( title = "House-Price-to-Income Ratio",
          caption = "Source: ACS 5-year, 2013-2017",
          fill = "Price-Income Ratio" ) +
    scale_fill_gradientn( colours=ocean.balance(10), guide = "colourbar" )


## LOGGED RATIO 

CenDF$log.price.ratio <- log( CenDF$HHInc_HousePrice_Ratio + 1 )

ggplot(CenDF) +
    geom_sf(aes(fill = log.price.ratio), color=NA) +
    coord_sf( datum=NA ) +
    labs( title = "House-Price-to-Income Ratio",
          caption = "Source: ACS 5-year, 2013-2017",
          fill = "Price-Income Ratio (log)" ) +
    scale_fill_gradientn( colours=ocean.balance(10), guide = "colourbar" )
```

