---
title: 'Saving Your Dorling Cartogram'
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: no
    css: 'lab-instructions.css'
    includes:
      after_body: 'footer.html'
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, eval=T)
```


<br>


---

<br>



You will be using your dorling cartogram file for you specific project city several times throughout the semester. 

You can re-create the cartogram steps each time, or else you can simply create it once and save the new shapefile on your computer for easy access. 

```{r, eval=F}
# load your census API key:
census_api_key("YOUR KEY GOES HERE")
```


```{r}

library( dplyr )
library( tidycensus )
library( sp )


phx.pop <-
get_acs( geography = "tract", variables = "B01003_001",
         state = "AZ", county = "Maricopa", geometry = FALSE ) %>% 
         select( GEOID, estimate ) %>%
         rename( POP=estimate )

phx.mhhi <- 
  get_acs( geography = "tract", variables = "B19013_001",
           state = "AZ", county = "Maricopa", geometry = FALSE ) %>% 
  select( GEOID, estimate ) %>% 
  rename( MHHI=estimate )


# get a census tract shapefile
# and add census data: 

library( tigris )
library( pander )

phx <- tracts( state="AZ", county="Maricopa", cb=TRUE, year=2015 )
phx <- merge( phx, phx.pop, by.x="GEOID", by.y="GEOID" )
phx <- merge( phx, phx.mhhi, by.x="GEOID", by.y="GEOID" )

head( phx@data ) %>% pander()

plot( phx )


library( rgdal )

# project map and remove empty tracts
phx <- spTransform( phx, CRS("+init=epsg:3395"))
phx <- phx[ phx$POP != 0 & (! is.na( phx$POP )) , ]


# devtools::install_github( "sjewo/cartogram" )

library( cartogram )  # spatial maps w/ tract size bias reduction
library( maptools )   # spatial object manipulation 
library( sf )         # 'simple features' flavor of shapefiles


# convert census tract polygons to dorling cartogram
# no idea why k=0.03 works, but it does - default is k=5
phx$pop.w <- phx$POP / 10000   # standardizes it to max of 1.5
phx_dorling <- cartogram_dorling( x=phx, weight="pop.w", k=0.03 )

plot( phx_dorling )



# install.packages( "tmap" )
library( tmap )       # thematic maps

tm_shape( phx_dorling ) + 
  tm_polygons( size="POP", col="MHHI", n=7, style="quantile", palette="Spectral" ) 



# WRITE TO FILE 

library( geojsonio )

phx_dorling <- spTransform( phx_dorling, CRS("+proj=longlat +datum=WGS84") )
geojson_write( phx_dorling, file="phx_dorling.geojson", geometry="polygon" )
```


Now you can load it from a local file, or upload it to GitHub so anyone can use the new dorling file: 

```{r}
library( geojsonio )
library( sp )
library( tmap ) 

# load from github
github.url <- "https://raw.githubusercontent.com/DS4PS/cpp-529-master/master/data/phx_dorling.geojson"
phx <- geojson_read( x=github.url,  what="sp" )

plot( phx )

# from local file path
# phx <- geojson_read( "data/phx_dorling.geojson", what="sp" )



phx <- spTransform( phx, CRS("+init=epsg:3395") )

bb <- st_bbox( c( xmin = -12519146, xmax = -12421368, 
                  ymax = 3965924, ymin = 3899074 ), 
               crs = st_crs("+init=epsg:3395"))

tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="MHHI", n=10, style="quantile", palette="Spectral" ) +
  tm_layout( "Dorling Cartogram", title.position=c("right","top") )

tmap_mode("view")
tm_basemap( "Stamen.Watercolor" ) +
  tm_shape( phx, bbox=bb ) + 
  tm_polygons( col="MHHI", n=7, style="quantile", palette="-inferno" ) 
```






<br>
<hr>
<br>

<style>
blockquote {
    padding: 11px 22px;
    margin: 0 0 22px;
    font-size: 18px;
    border-left: 5px solid lightgray;
}

</style>

